---
title: "An introduction to hdsp package with examples."
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE}
library(hdsp)
```

## Overview
This package is implementation of various shrinkage techniques with application to the so-called expected utility portfolios. It contains three main functions: EUShrinkPortfolio, CovarEstim and MeanEstim. 

The following setup will be used throughout all examples:
```{r experiment setup}

n<-3e2 # number of realizations
p<-.5*n # number of assets
gamma<-1

# the target for covariance shrinkage
TM <- matrix(0, p, p) 
diag(TM) <- 1

# the target for shrinkage of the mean vector
mu_0 <- rep(0, p)
mu_0[1:10] <- 10:1
```

All data matrices in the package are handled in a way that the variables are in rows and observations- in columns.
```{r}
x <- matrix(data = rnorm(n*p), nrow = p, ncol = n) # asset returns
```

## EU Portfolios
The general function performing portfolio creation is EUShrinkPortfolio. Essentially, it is a function dispatcher which chooses a method according to arguments type and subtype. It returns either a portfolio in the form of ExUtil_portfolio S3 class or a human-readable error message. Argument type represents the type of shrinkage to use, while subtype is a particular method. EUShrinkPortfolio should be used in interactive programming as it is more user-friendly and slower than underlying methods, while methods themselves should be exploited for writing more complicated but at the same time efficient and fast code.

### EU portfolio with shrinked weights

Weight-shrinked portfolio from [IEEE2020] is constructed by setting type='weights'. Subtype argument is not set since the method is the only one of its type implemented.
```{r}
portfolio <- EUShrinkPortfolio(x=x, gamma=gamma, type='weights', b=rep(1/p,p))
str(portfolio)
```

### EU portfolio with shrinked means

In this kind of portfolios shrinkage for mean values is used with no shrinkage of other elements. The subtype options available are: James-Stein and Bayes-Stein from [Jorion 1986] and Bodnar-Okhrin-Parolya from [BOP19].

Bodnar-Okhrin-Parolya [BOP19]:
```{r}
portfolio <- EUShrinkPortfolio(x=x, gamma=gamma, type='mean', 
                               subtype='BOP19', mu_0=mu_0)
str(portfolio)
```

Bayes-Stein [Jorion 1986]:
```{r}
portfolio <- EUShrinkPortfolio(x=x, gamma=gamma, type='mean', 
                               subtype='Bayes-Stein', mu_0=mu_0)
str(portfolio)
```

James-Stein [Jorion 1986]:
```{r}
portfolio <- EUShrinkPortfolio(x=x, gamma=gamma, type='mean', 
                               subtype='James-Stein', mu_0=mu_0)
str(portfolio)
```

### EU portfolio with a shrinked covariance matrix

In this type (type='cov') the covariance matrix is obtained via shrinkage estimators. The available subtypes are: 'BGP14' and 'LW20' corresponding to methods from [BGP14] and [LW20].

```{r}
portfolio <- EUShrinkPortfolio(x=x, gamma=gamma, type='cov',
                               subtype='BGP14', TM=TM)
str(portfolio)
```


## Shrinkage estimators for covariance matrix
The next important function is CovarEstim. This is a dispatcher for functions computing the covariance matrix of the data.
```{r}
Mtrx_naive <- CovarEstim(x, type="naive")
# Mtrx_naive[1:6,1:6]
format(Mtrx_naive[1:6,1:6], scientific = TRUE, digits = 3)
```

```{r}
Mtrx_bgp <- CovarEstim(x, type="BGP14", TM=TM, SCM=Mtrx_naive)
# Mtrx_bgp[1:6,1:6]
format(Mtrx_bgp[1:6,1:6], scientific = TRUE, digits = 3)
```


## Shrinkage estimators for mean vectors

```{r}
Mean_BOP <- MeanEstim(x, type="BOP19", mu_0=mu_0)
head(Mean_BOP)
```

## Custom Expected Utility portfolios
There is a possibility to make custom EU portfolios having mean values of returns and an inverse covariance matrix.
```{r}
cov_mtrx <- CovarEstim(x, type="LW20", TM=TM)
invSS <- solve(cov_mtrx)
means <- MeanEstim(x, type="BOP19", mu_0=mu_0)

cust_port_BS_LW <- new_ExUtil_portfolio_custom(mean_vec=means, 
                                               inv_cov_mtrx=invSS, gamma=2)
str(cust_port_BS_LW)
```
